---
title: "Asignment2B"
author: "Cecilia De La Fuente Chavez"
format: html
execute:
  echo: true
editor: visual
---

## Approach Deliverable

### Overview

Classification models are new to me and I only learned the basics in class yesterday, I will approach this assignment by first examining the values and compare the resulting confusion matrices, and calculate the requested metrics. One challenge I anticipate is making sure I correctly handle these calculations.

## Code Base Deliverable

### Load library

```{r}
library(tidyverse)
```

### Read data from GitHub

```{r}
url <- "https://raw.githubusercontent.com/acatlin/data/refs/heads/master/penguin_predictions.csv"

df <- read_csv(
  file = url,
  show_col_types = FALSE,
  progress = FALSE
)

head(df)
```

### Step 1

Get count of penguins by actual sex, total count, and percent of each sex

```{r}
penguin_count_sex <- df |>
  count(sex) |>
  mutate(total_rows = sum(n),
         sex_percent = n/total_rows,
         )

print(penguin_count_sex)
```

Plot the distribution

```{r}
ggplot(penguin_count_sex, aes(x = sex, y = n, fill = sex)) +
  geom_col() +
  labs(
    title = "Penguin Sex Distribution"
  ) +
  geom_text(
    aes(label = paste0(round(sex_percent * 100, 1), "%")),
    vjust = -0.3
  )
```

### Step 2

Get c
