---
title: "Asignment2B"
author: "Cecilia De La Fuente Chavez"
format: html
execute:
  echo: true
editor: visual
---

## Approach Deliverable

### Overview

Classification models are new to me and I only learned the basics in class yesterday, I will approach this assignment by first examining the values and compare the resulting confusion matrices, and calculate the requested metrics. One challenge I anticipate is making sure I correctly handle these calculations.

## Code Base Deliverable

### Load library

```{r}
library(tidyverse)
```

### Read data from GitHub

```{r}
url <- "https://raw.githubusercontent.com/acatlin/data/refs/heads/master/penguin_predictions.csv"

df <- read_csv(
  file = url,
  show_col_types = FALSE,
  progress = FALSE
)

head(df)
```

Get count of penguins by actual sex, total count, and percent of each sex

```{r}
penguin_dist <- df |>
  count(sex) |>
  mutate(total_rows = sum(n),
         sex_percent = n/total_rows,
         )

print(penguin_dist)
```

Plot the distribution

```{r}
ggplot(penguin_dist, aes(x = sex, y = n, fill = sex)) +
  geom_col() +
  labs(
    title = "Penguin Sex Distribution"
  ) +
  geom_text(
    aes(label = paste0(round(sex_percent * 100, 1), "%")),
    vjust = -0.3
  )
```

Null metrics calculations

```{r}
null_metrics <- penguin_dist |>
  summarize(
    accuracy = max(sex_percent),
    error_rate = 1 - max(sex_percent)
  )

print(null_metrics)
```

Recompute predicted class

```{r}
prediction_results <- df |>
  mutate(
    pred_02 = ifelse(.pred_female > 0.2, 1, 0),
    pred_05 = ifelse(.pred_female > 0.5, 1, 0),
    pred_08 = ifelse(.pred_female > 0.8, 1, 0)
  )

print(prediction_results)
```

Metrics for .5

```{r}
prediction_results <- prediction_results |>
  mutate(
    TP = sex == "female" & pred_05 == 1,
    FP = sex == "male" & pred_05 == 1,
    TN = sex == "male" & pred_05 == 0,
    FN = sex == "female" & pred_05 == 0
  )

metrics_05 <- prediction_results |>
  summarize(
    TP = sum(TP),
    FP = sum(FP),
    TN = sum(TN),
    FN = sum(FN)
  ) |>
  mutate(
    accuracy = (TP + TN) / (TP + FP + TN + FN),
    precision = TP / (TP + FP),
    recall = TP / (TP + FN),
    f1 = 2 * precision * recall / (precision + recall)
  )

print(metrics_05)
```

Metrics for .2

```{r}
prediction_results <- prediction_results |>
  mutate(
    TP = sex == "female" & pred_02 == 1,
    FP = sex == "male" & pred_02 == 1,
    TN = sex == "male" & pred_02 == 0,
    FN = sex == "female" & pred_02 == 0
  )

metrics_02 <- prediction_results |>
  summarize(
    TP = sum(TP),
    FP = sum(FP),
    TN = sum(TN),
    FN = sum(FN)
  ) |>
  mutate(
    accuracy = (TP + TN) / (TP + FP + TN + FN),
    precision = TP / (TP + FP),
    recall = TP / (TP + FN),
    f1 = 2 * precision * recall / (precision + recall)
  )

print(metrics_02)
```

Metrics for .8

```{r}
prediction_results <- prediction_results |>
  mutate(
    TP = sex == "female" & pred_08 == 1,
    FP = sex == "male" & pred_08 == 1,
    TN = sex == "male" & pred_08 == 0,
    FN = sex == "female" & pred_08 == 0
  )

metrics_08 <- prediction_results |>
  summarize(
    TP = sum(TP),
    FP = sum(FP),
    TN = sum(TN),
    FN = sum(FN)
  ) |>
  mutate(
    accuracy = (TP + TN) / (TP + FP + TN + FN),
    precision = TP / (TP + FP),
    recall = TP / (TP + FN),
    f1 = 2 * precision * recall / (precision + recall)
  )

print(metrics_08)
```

## Conclusion

The different probability thresholds affected the results. At 0.2 , the model had the highest recall (94.9%), correctly identifying the largest proportion of actual females, but it also had more false positives and lower precision (86.0%). At 0.5, precision and recall were both 92.3%. At 0.8, the model had the highest accuracy (94.6%), precision (94.7%), and F1 score (93.5%). Overall, the 0.8 threshold provided the best overall performance.
